---
title: "ECOL 610: NEON Assignment 1"
author: "Group - Santa Rita Experimental Range (SRER)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
header-includes:
  - \usepackage{caption}
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'ECOL610_NEONAssignment1_SRER.pdf')) 
  })
---
  - Emily Swartz
  - Shahriar Shah Heydari
  - Stephanie Cardinalli
  - George Woolsey

# In Class

## Setup
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  , fig.height = 5
  , fig.width = 7
)
```

This is the R script for ECOL610 NEON assignment 1
This script will provide code to carry out assignment 1 for the Central Plains Experimental Range - you will need to modify this code to carry it out for your site.

First, load in the needed packages. Install the packages if needed.

```{r}
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)
library(latex2exp)
```


Load in the data. You will need to change the file path for your directory.

```{r}
#load in the data
CPER_30 <- read.csv("../data/Central Plains Experimental Range - 30 min.csv")
CPER_daily <- read.csv("../data/Central Plains Experimental Range - daily.csv")
```

## Question 1

Calculate GPP and compare to NEON GPP for summer and winter week

*NEE is defined, by convention, as $CO_{2}$ flux from the ecosystem to the atmosphere. It corresponds to a negative carbon input to ecosystems. NEE is defined in this way because atmospheric scientists, who originated the term, seek to document net sources of $CO_{2}$ to the atmosphere (i.e., NEE) that account for rising atmospheric $CO_{2}$ concentration. Therefore, $CO_{2}$ input to the ecosystem is a negative NEE.*

[Chapin, F. S., Matson, P. A., Mooney, H. A., & Vitousek, P. M. (2002). Principles of terrestrial ecosystem ecology.](https://d1wqtxts1xzle7.cloudfront.net/55690956/Principles_of_terrestrial_ecosystem_ecology-with-cover-page-v2.pdf?Expires=1663106506&Signature=ZLKRpouXVl6Q2oVAvMbYfcyWZT227z~A0BOTNMvx3nr-hzPv-aQr2DF-vvK~O2T8ygmVtbYXdNlXfNAE8FYZ70B2OOHPU8HHIhXPwKW90Mf~SYyj2xIQBIb0gMK4mZ6lJLG~eO7cPoLuK974yvVy5zdcnJt81MhsSB2vPb3w8l-QijHyNlYmpr43FYR50UuYNAib58kuaUNYxN-jMFaLVLS6fvYxV93ToeH3mILBD3mMliAUAViXzXngzVVuLQXXyJodsR1JbR54PJ-Uhyeitj7PI9Qq1Rtpz1Y0gRIkXd5DiJenTOXLTpc1jD~OYBqyGowRjcSPMSgyMT1cilxGQQ__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA) p.208

$$
\begin{aligned}
NEE = R_{E} - GPP
\end{aligned}
$$
$$
\begin{aligned}
GPP = R_{E} - NEE
\end{aligned}
$$

```{r}
#calculate GPP from NEE and Re using equation NEE = Re - GPP (negative values land C storage)
CPER_30 <- CPER_30 %>% 
  dplyr::mutate(
    GPP_calc = as.numeric(RE - NEE)
    , GPP_is_equal = round(GPP_calc, 3) == round(GPP, 3)
  )

# are there rows where the data is not equal?
  nrow(
    CPER_30 %>% 
      dplyr::filter(
        GPP_is_equal == FALSE
        & !is.na(GPP)
      )
  )


#clipping the data to a winter and summer week
 #choose your own weeks (without data gaps) for your site!
#February 16-22, 2020
CPER_winter_week <- filter(CPER_30, Year == "2020" & DOY == 47:53)
#June 21-27, 2020
CPER_summer_week <- filter(CPER_30, Year == "2020" & DOY == 173:180)

#creating long format for plotting
CWW_stacked <-gather(CPER_winter_week, "GPP_type", "GPP_value", c(9,15))
CSW_stacked <-gather(CPER_summer_week, "GPP_type", "GPP_value", c(9,15))

#winter
#plot of NEON vs calculated GPP
ggplot(CWW_stacked,aes(x=DOY.total,y=GPP_value, color = GPP_type)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(expression(paste("GPP ( mol CO"[2]*" m"^{-2}, "day"^{-2}*")"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Winter GPP Comparison")

#summer
#plot of NEON vs calculated GPP
ggplot(CSW_stacked,aes(x=DOY.total,y=GPP_value, color = GPP_type)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(latex2exp::TeX("$GPP \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Summer GPP Comparison")

```

## Question 2 

annual patterns in C exchange and environmental properties

```{r}
#reduce data to just 2020
CPER_2020 <- filter(CPER_daily, Year == "2020")
#plot of NEE
ggplot(CPER_2020,aes(x=DOY,y=NEE)) + geom_point() + 
    xlab("Day of Year (2020)") + 
    ylab(latex2exp::TeX("$NEE \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    labs(title = "Annual NEE",
         subtitle = "CPER")
#plot of GPP
ggplot(CPER_2020,aes(x=DOY,y=GPP)) + geom_point() + 
    xlab("Day of Year (2020)") + 
    ylab(latex2exp::TeX("$GPP \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    labs(title = "Annual GPP",
         subtitle = "CPER")
#plot of Re
ggplot(CPER_2020,aes(x=DOY,y=RE)) + geom_point() + 
    xlab("Day of Year (2020)") + 
    ylab(latex2exp::TeX("$R_{E} \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    labs(title = latex2exp::TeX("$Annual \\; R_{E}$"),
         subtitle = "CPER")
#plot of soil temp
ggplot(CPER_2020,aes(x=DOY,y=TS)) + geom_point() + 
    xlab("Day of Year (2020)") + 
    ylab("Soil temperature (\u00B0C)") +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    labs(title = "Annual Soil Temperature",
         subtitle = "CPER")
#plot of air temp
ggplot(CPER_2020,aes(x=DOY,y=TA)) + geom_point() + 
    xlab("Day of Year (2020)") + 
    ylab("Air temperature (\u00B0C)") +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    labs(title = "Annual Air Temperature",
         subtitle = "CPER")
#plot of soil moisture
ggplot(CPER_2020,aes(x=DOY,y=SWC)) + geom_point() + 
    xlab("Day of Year (2020)") + 
    ylab("Soil Water Content (%)") +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    labs(title = "Annual Soil Water Content",
         subtitle = "CPER")

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Assignment

## Load data
```{r}
remove(list=ls())
# what is your site name?
site <- "Santa Rita Experimental Range"
# load 30 min data
f_list <- list.files(path = "../data/", pattern="*30 min.csv")
for (i in 1:length(f_list)){
  nm <- stringr::word(f_list[i], start = 1, sep = "30 min") %>% 
    stringr::str_replace_all(pattern = "[[:punct:]]", replacement = "") %>% 
    stringr::str_trim() %>% 
    stringr::str_squish()
  temp <- read.csv(paste0("../data/", f_list[i])) %>% 
    dplyr::mutate(neon_site_name = nm) %>% 
    dplyr::relocate(neon_site_name) %>% 
    dplyr::rename_with(~ tolower(
      gsub(" ", "_", 
         str_trim(gsub("\\s+", " ", .x))
      )
    ))
  if(i==1){
    dta_30min <- temp
  }else{
    dta_30min <- dplyr::union_all(dta_30min, temp)
  }
  remove(temp)
}
# load daily data
f_list <- list.files(path = "../data/", pattern="*daily.csv")
for (i in 1:length(f_list)){
  nm <- stringr::word(f_list[i], start = 1, sep = "daily") %>% 
    stringr::str_replace_all(pattern = "[[:punct:]]", replacement = "") %>% 
    stringr::str_trim() %>% 
    stringr::str_squish()
  temp <- read.csv(paste0("../data/", f_list[i])) %>% 
    dplyr::mutate(neon_site_name = nm) %>% 
    dplyr::relocate(neon_site_name) %>% 
    dplyr::rename_with(~ tolower(
      gsub(" ", "_", 
         str_trim(gsub("\\s+", " ", .x))
      )
    ))
  if(i==1){
    dta_1day <- temp
  }else{
    dta_1day <- dplyr::union_all(dta_1day, temp)
  }
  remove(temp)
}
# create dates and record counts
# day
dta_1day <- dta_1day %>% 
  dplyr::mutate(
    date_id = lubridate::make_date(year = year, month = month, day = day)
    , week = lubridate::week(date_id)
    , has_gpp = ifelse(!is.na(gpp), 1, 0)
    , season =
        dplyr::case_when(
          month %in% c(1:3, 12) ~ "Winter"
          , month %in% c(6:8) ~ "Summer"
          , TRUE  ~ "other")
  ) %>% 
  dplyr::group_by(neon_site_name, week, year) %>% 
  dplyr::mutate(is_full_week = sum(has_gpp)==7) %>% 
  dplyr::ungroup()
# 30
dta_30min <- dta_30min %>% 
  dplyr::mutate(
    date_id = lubridate::make_date(year = year, month = month, day = day)
    , time_id = lubridate::make_datetime(year = year, month = month, day = day
                        , hour = floor(hour)
                        , min = (hour-floor(hour))*60
                )
    , week = lubridate::week(date_id)
    , has_gpp = ifelse(!is.na(gpp), 1, 0)
    , season =
        dplyr::case_when(
          month %in% c(1:3, 12) ~ "Winter"
          , month %in% c(6:8) ~ "Summer"
          , TRUE  ~ "other")
  ) %>% 
  dplyr::group_by(neon_site_name, week, year) %>% 
  dplyr::mutate(is_full_week = sum(has_gpp)==24*2*7) %>% 
  dplyr::ungroup()
```

The draft questions for that assignment are:

## Question 1 

Using the 30 minute data. Calculate GPP for your site. 

```{r}
dta_30min <- dta_30min %>% 
  dplyr::mutate(
    gpp_calc = as.numeric(re - nee)
    , gpp_is_equal = round(gpp_calc, 2) == round(gpp, 2)
  )
# quick plot of gpp equality
dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & !is.na(gpp)
  ) %>% 
ggplot(., aes(x = gpp_is_equal)) +
  geom_bar(
    aes(fill = gpp_is_equal)
    , width = 0.7
  ) +
  geom_text(
    aes(label = scales::comma(..count..))
    , stat = "count"
    , vjust = -0.5
    , color = "black"
  ) +
  scale_fill_viridis_d(alpha = 0.8) +
  scale_y_continuous(labels = scales::comma) +
  xlab("GPP Calc. = GPP? (accuracy = .01)") +
  labs(
    title = "GPP Calc. = GPP? (accuracy = .01)"
    , subtitle = paste0(site, " (30-min data)")
  ) +
  theme_bw() +
  theme(
    legend.position = "none"
  )
```

Choose a winter and summer week in your dataset. Create a plot with both your calculated GPP and the NEON GPP. 

```{r}
# select week with full data for winter and summer
sum_wk <- dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & season == "Summer"
    & year == 2020
    & is_full_week == TRUE
  ) %>% 
  dplyr::slice_sample(n = 1)
wint_wk <- dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & season == "Winter"
    & year == 2020
    & is_full_week == TRUE
  ) %>% 
  dplyr::slice_sample(n = 1)

# plot winter week
dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & year == 2020
    & week == wint_wk$week[1]
  ) %>% 
ggplot(., aes(x = time_id)) +
  geom_point(
    aes(y = gpp, color = "GPP")
    , size = 2.5
  ) +
  geom_point(
    aes(y = gpp_calc, color = "GPP Calc.")
    , size = 0.9
  ) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  xlab("Day of Year") + 
  ylab(latex2exp::TeX("$GPP \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
  labs(
    title = "Winter Week Sample: Calculated GPP vs. NEON GPP"
    , subtitle = paste0(site, " (30-min data)")
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))

# plot summer week
dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & year == 2020
    & week == sum_wk$week[1]
  ) %>% 
ggplot(., aes(x = time_id)) +
  geom_point(
    aes(y = gpp, color = "GPP")
    , size = 2.5
  ) +
  geom_point(
    aes(y = gpp_calc, color = "GPP Calc.")
    , size = 0.9
  ) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  xlab("Day of Year") + 
  ylab(latex2exp::TeX("$GPP \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
  labs(
    title = "Summer Week Sample: Calculated GPP vs. NEON GPP"
    , subtitle = paste0(site, " (30-min data)")
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))
```


### a. 

How do the calculated and NEON GPP's compare for your site? Why are they the same or different? (Note: NEON uses eddy covariance to calculate these metrics - more about that [here](https://youtu.be/CR4Anc8Mkas)).

\textcolor{blue}{The majority of data points have a calculated GPP ($GPP = R_{E} - NEE$) equal to the NEON GPP. There are `r scales::comma(dta_30min %>% dplyr::filter(neon_site_name == site & gpp_is_equal == TRUE) %>% nrow())` records for which the GPP is equal (accuracy = 0.01) and `r scales::comma(dta_30min %>% dplyr::filter(neon_site_name == site & gpp_is_equal == FALSE) %>% nrow())` records for which the GPP is not equal. In cases where the GPP is not equal, the difference is negligible and likely due to rounding due to number accuracy.}


### b. 

How do your sites GPP values vary between summer and winter? What do you think is driving these differences?

```{r}
# plot points
dta_30min %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
    & season %in% c("Summer", "Winter")
  ) %>% 
ggplot(., aes(x = time_id, y = gpp, color = season)) +
  geom_point(
    # size = 0.2
    alpha = 0.4
    , shape = "."
  ) +
  geom_smooth(
    method = "lm"
    , formula=y~1
    , se = FALSE
    , lwd = 1.1
    , linetype = "solid"
    , alpha = 1
  ) +
  facet_wrap(~neon_site_name) +
  scale_color_viridis_d(option = "turbo", direction = -1) +
  xlab("Day of Year") + 
  ylab(latex2exp::TeX("$GPP \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
  labs(
    title = "2020 NEON GPP (30-min data)"
    , subtitle = paste0(site)
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
    , axis.text.x = element_text(angle = 60, hjust=1)
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))



```


\textcolor{blue}{Based on 2020 data, The average 30-minute GPP for the NEON site `r site` is higher in the winter months than in the summer months. This is opposite of the pattern for GPP seen at the CPER site. At the `r site`, this pattern is likely driven by low summer precipitation and high summer temperatures to the winter.}

### c. 

CPER is a semi-arid grassland. Compare GPP values between your site and CPER - why might they be different or similar?

```{r}
# plot points
dta_30min %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
    & season %in% c("Summer", "Winter")
  ) %>% 
  dplyr::group_by(neon_site_name, season, year) %>% 
  dplyr::summarise(mean_gpp = mean(gpp, na.rm = TRUE)) %>% 
ggplot(., aes(x = season, y = mean_gpp, fill = season)) +
  geom_col(
    width = 0.7
    , alpha = 0.8
  ) +
  geom_text(
    aes(label = scales::comma(mean_gpp, accuracy = 0.01))
    , vjust = -0.3
    , color = "black"
  ) +
  facet_wrap(~neon_site_name) +
  scale_fill_viridis_d(option = "turbo", direction = -1) +
  xlab("Season") + 
  ylab(latex2exp::TeX("Mean 30-min $GPP \\; ( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-2})$")) +
  labs(
    title = "2020 Seasonal NEON GPP (30-min data)"
    , subtitle = paste0(site)
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))

```

\textcolor{blue}{Based on 2020 data, The average 30-minute GPP for the NEON site `r site` is higher in the winter months (`r scales::comma(mean((dta_30min %>% dplyr::filter(neon_site_name == site & year == 2020 & season %in% c("Winter")))$gpp), accuracy = 0.01)`) than in the summer months (`r scales::comma(mean((dta_30min %>% dplyr::filter(neon_site_name == site & year == 2020 & season %in% c("Summer")))$gpp), accuracy = 0.01)`). This is opposite of the pattern for GPP seen at the CPER site. At the `r site`, this pattern is likely driven by low summer precipitation and high summer temperatures to the winter.}


## Question 2 

Using the daily data. Select a single year of data for your site. Plot NEE, GPP, Re, soil temperature, air temperature, and soil water content against Day of Year.

```{r}
p_fn <- function(my_var, my_year) {
  my_title <- paste0(toupper(my_var), " - ", as.character(my_year))
  #plot
  (
    dta_1day %>% 
    dplyr::filter(
      year == my_year
    ) %>% 
    ggplot(., aes_string(x = "date_id", y = my_var, color = "neon_site_name")) +
      geom_point() +
      xlab("Date") + 
      ylab(toupper(my_var)) +
      labs(
        title = my_title
        # , subtitle = site
      ) +
      scale_fill_brewer(type = "qual", palette = "Dark2", direction = -1) +
      theme_bw() +
      theme(
        legend.position = "top"
        , legend.direction = "horizontal"
        , legend.title = element_blank()
      ) +
      guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))
  )
  
}
# vars
v <- c("nee", "gpp", "re", "ts", "ta", "swc")
for (i in 1:length(v)) {
  print( p_fn(my_var = v[i], my_year = 2020) )
}

```


### a. 

Describe the annual patterns in each plot and what you think drives them.

### b. 

Compare annual values of NEE at your site vs. CPER. Which exchanges more carbon? Why do you think that is?

