---
title: "ECOL 610: NEON Assignment 1"
author: "Group - Santa Rita Experimental Range (SRER)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
linkcolor: blue
header-includes:
  - \usepackage{caption}
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'ECOL610_NEONAssignment1_SRER.pdf')) 
  })
---

# Group Members

  - Emily Swartz
  - Shahriar Shah Heydari
  - Stephanie Cardinalli
  - George Woolsey


# Setup
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  , fig.height = 5
  , fig.width = 7
)
```

First, load in the needed packages. Install the packages if needed.

```{r}
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)
library(latex2exp)
library(psych)
remove(list=ls())
```

# Introduction

*Net ecosystem exchange (NEE) is defined, by convention, as $CO_{2}$ flux from the ecosystem to the atmosphere. It corresponds to a negative carbon input to ecosystems. NEE is defined in this way because atmospheric scientists, who originated the term, seek to document net sources of $CO_{2}$ to the atmosphere (i.e., NEE) that account for rising atmospheric $CO_{2}$ concentration. Therefore, $CO_{2}$ input to the ecosystem is a negative NEE.*

[Chapin, F. S., Matson, P. A., Mooney, H. A., & Vitousek, P. M. (2002). Principles of terrestrial ecosystem ecology.](https://d1wqtxts1xzle7.cloudfront.net/55690956/Principles_of_terrestrial_ecosystem_ecology-with-cover-page-v2.pdf?Expires=1663106506&Signature=ZLKRpouXVl6Q2oVAvMbYfcyWZT227z~A0BOTNMvx3nr-hzPv-aQr2DF-vvK~O2T8ygmVtbYXdNlXfNAE8FYZ70B2OOHPU8HHIhXPwKW90Mf~SYyj2xIQBIb0gMK4mZ6lJLG~eO7cPoLuK974yvVy5zdcnJt81MhsSB2vPb3w8l-QijHyNlYmpr43FYR50UuYNAib58kuaUNYxN-jMFaLVLS6fvYxV93ToeH3mILBD3mMliAUAViXzXngzVVuLQXXyJodsR1JbR54PJ-Uhyeitj7PI9Qq1Rtpz1Y0gRIkXd5DiJenTOXLTpc1jD~OYBqyGowRjcSPMSgyMT1cilxGQQ__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA) p.208

Key point: **$CO_{2}$ input to the ecosystem is a negative NEE**

$$
\begin{aligned}
NEE = R_{E} - GPP
\end{aligned}
$$

$$
\begin{aligned}
GPP = R_{E} - NEE
\end{aligned}
$$

$$
\begin{aligned}
NEP = GPP - R_{E}
\end{aligned}
$$

# Load Data

```{r}
# what is your site name?
site <- "Santa Rita Experimental Range"
# load 30 min data 
  # I put both CPER and SRER data in this directory
  # all data will be loaded in the same R data set
f_list <- list.files(path = "../data/", pattern="*30 min.csv")
for (i in 1:length(f_list)){
  nm <- stringr::word(f_list[i], start = 1, sep = "30 min") %>% 
    stringr::str_replace_all(pattern = "[[:punct:]]", replacement = "") %>% 
    stringr::str_trim() %>% 
    stringr::str_squish()
  temp <- read.csv(paste0("../data/", f_list[i])) %>% 
    dplyr::mutate(neon_site_name = nm) %>% 
    dplyr::relocate(neon_site_name) %>% 
    dplyr::rename_with(~ tolower(
      gsub(" ", "_", 
         str_trim(gsub("\\s+", " ", .x))
      )
    ))
  if(i==1){
    dta_30min <- temp
  }else{
    dta_30min <- dplyr::union_all(dta_30min, temp)
  }
  remove(temp)
}
# load daily data
f_list <- list.files(path = "../data/", pattern="*daily.csv")
for (i in 1:length(f_list)){
  nm <- stringr::word(f_list[i], start = 1, sep = "daily") %>% 
    stringr::str_replace_all(pattern = "[[:punct:]]", replacement = "") %>% 
    stringr::str_trim() %>% 
    stringr::str_squish()
  temp <- read.csv(paste0("../data/", f_list[i])) %>% 
    dplyr::mutate(neon_site_name = nm) %>% 
    dplyr::relocate(neon_site_name) %>% 
    dplyr::rename_with(~ tolower(
      gsub(" ", "_", 
         str_trim(gsub("\\s+", " ", .x))
      )
    ))
  if(i==1){
    dta_1day <- temp
  }else{
    dta_1day <- dplyr::union_all(dta_1day, temp)
  }
  remove(temp)
}
# create dates and record counts
# 1-day
dta_1day <- dta_1day %>% 
  dplyr::mutate(
    date_id = lubridate::make_date(year = year, month = month, day = day)
    , week = lubridate::week(date_id)
    , has_gpp = ifelse(!is.na(gpp), 1, 0)
    , season =
        dplyr::case_when(
          month %in% c(1:3, 12) ~ "Winter"
          , month %in% c(6:8) ~ "Summer"
          , TRUE  ~ "Other")
  ) %>% 
  dplyr::group_by(neon_site_name, week, year) %>% 
  dplyr::mutate(is_full_week = sum(has_gpp)==7) %>% 
  dplyr::ungroup()
# 30-min
dta_30min <- dta_30min %>% 
  dplyr::mutate(
    date_id = lubridate::make_date(year = year, month = month, day = day)
    , time_id = lubridate::make_datetime(year = year, month = month, day = day
                        , hour = floor(hour)
                        , min = (hour-floor(hour))*60
                )
    , week = lubridate::week(date_id)
    , has_gpp = ifelse(!is.na(gpp), 1, 0)
    , season =
        dplyr::case_when(
          month %in% c(1:3, 12) ~ "Winter"
          , month %in% c(6:8) ~ "Summer"
          , TRUE  ~ "other")
  ) %>% 
  dplyr::group_by(neon_site_name, week, year) %>% 
  dplyr::mutate(is_full_week = sum(has_gpp)==24*2*7) %>% 
  dplyr::ungroup()
```

# Assignment

Make some plots for class on September 20, 2022. For these plots, please use the CPER daily data for 2020 from DOY 175 to 225. Make four plots:

  * NEE, GPP, Re (all on y-axis) vs soil water content (x-axis)
  * NEE, GPP, Re (all on y-axis) vs air temperature (x-axis)
  * NEE, GPP, Re (all on y-axis) vs soil temperature (x-axis)
  * NEE, GPP, Re (all on y-axis) vs photon flux density (x-axis)

Remember there is more info about these variables in the [metadata file](https://colostate-my.sharepoint.com/:x:/g/personal/gwoolsey_colostate_edu/EUC3O5JNO8REvf7EnuAXqZ8BoA7NWmOF6FE2g_e4ycz6sg?e=oZvoRv)

## Example plot for 1 dependent variable & 1 independent variable

```{r}
dta_1day %>% 
    dplyr::filter(
      year == 2020
      & doy %in% c(175:225)
      & neon_site_name %in% c("Central Plains Experimental Range")
    ) %>%
  ggplot(., aes(x = swc, y = nee)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Example NEE vs. SWC"
    , subtitle = "Central Plains Experimental Range"
    , caption = "2020 - DOY 175 to 225"
  ) +
  theme_bw()
```

## Example plot for 3 dependent variables & 1 independent variable

```{r}
dta_1day %>% 
    dplyr::filter(
      year == 2020
      & doy %in% c(175:225)
      & neon_site_name %in% c("Central Plains Experimental Range")
    ) %>%
  ggplot(.) + 
  geom_point(aes(x = swc, y = nee, color = "NEE")) +
  geom_point(aes(x = swc, y = gpp, color = "GPP")) +
  geom_point(aes(x = swc, y = re, color = "Re")) +
  geom_smooth(aes(x = swc, y = nee, color = "NEE"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = swc, y = gpp, color = "GPP"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = swc, y = re, color = "Re"), method = "lm", se = FALSE) +
  labs(
    title = "Example NEE, GPP, Re vs. SWC"
    , subtitle = "Central Plains Experimental Range"
    , caption = "2020 - DOY 175 to 225"
  ) +
  theme_bw()
```

## Define Function for Extendable Plotting

```{r}
# define dependent vars
my_plot_function <- function(
  dep_vars = c("gpp", "nee", "re")
  , dep_vars_lab = c(latex2exp::TeX("$GPP$"), latex2exp::TeX("$NEE$"), latex2exp::TeX("$R_E$"))
  , st_doy = 175
  , end_doy = 225
  , yr = 2020
  , my_cov = "swc"
  , my_cov_lab = "SWC"
  , sites = c("Central Plains Experimental Range")
){
  (
  # filter data
    dta_1day %>% 
      dplyr::filter(
        year == yr
        & doy %in% c(st_doy:end_doy)
        & neon_site_name %in% sites
      ) %>%
  # pivot dependent vars
    tidyr::pivot_longer(
      cols = tidyselect::all_of(dep_vars)
      , names_to = "dependent_var_name"
      , values_to = "dependent_var_value"
      , values_drop_na = FALSE
    ) %>% 
    dplyr::arrange(
      neon_site_name
      , date_id
      , dependent_var_name
    ) %>% 
  # plot
    ggplot(
      data = . # "." means the data that is passed to ggplot via the pipe "%>%"
      , mapping = aes_string(
        x = my_cov
        , y = "dependent_var_value"
        , color = "dependent_var_name"
      )
    ) +
    geom_point(
      alpha = 0.5
      , size = 0.8
    ) +
    geom_smooth(
      method = "lm"
      , se = FALSE
      , lwd = 1
      , alpha = 0.8
    ) + 
    facet_wrap(~neon_site_name, scales = "free") +
    scale_color_brewer(
      type = "qual"
      , palette = "Paired"
      , direction = -1
      , labels = dep_vars_lab
    ) +
    xlab(my_cov_lab) + 
    ylab(latex2exp::TeX("$mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1}$")) +
    labs(
      title = paste0(yr, " NEON Site Carbon Fluxes (1-day data)")
      , subtitle = my_cov_lab
      , caption = paste0(
          as.Date(st_doy-1, origin = paste0(yr,"-01-01"))
          , " to "
          , as.Date(end_doy-1, origin = paste0(yr,"-01-01"))
      )
    ) +
    theme_bw() +
    theme(
      legend.position = "top"
      , legend.direction = "horizontal"
      , legend.title = element_blank()
    ) +
    guides(color = guide_legend(override.aes = list(size = 5)))
  )
}
```

## All covariates 2020 DOY 175-225

```{r, fig.height = 4, fig.width = 7}
v <- c("swc", "ta", "ts", "ppfd_in")
t <- c(
  "Soil Water Content (%)"
  , "Air temperature (\u00B0C)"
  , "Soil Temperature (\u00B0C)"
  , latex2exp::TeX("$\\mu$mol photons $m^{-2} \\cdot s^{-1}$")
)
for (i in 1:length(v)) {
  print(
    my_plot_function(
      dep_vars = c("gpp", "nee", "re")
      , dep_vars_lab = c(latex2exp::TeX("$GPP$"), latex2exp::TeX("$NEE$"), latex2exp::TeX("$R_E$"))
      , st_doy = 175
      , end_doy = 225
      , yr = 2020
      , my_cov = v[i]
      , my_cov_lab = t[i]
      , sites = c("Central Plains Experimental Range")
    )
  )
}
```

## Compare CPER to SRER 2020 full year

```{r, fig.height = 4, fig.width = 7}
v <- c("swc", "ta", "ts", "ppfd_in")
t <- c(
  "Soil Water Content (%)"
  , "Air temperature (\u00B0C)"
  , "Soil Temperature (\u00B0C)"
  , latex2exp::TeX("PPFD ($\\mu$mol photons $m^{-2} \\cdot s^{-1}$)")
)
for (i in 1:length(v)) {
  print(
    my_plot_function(
      dep_vars = c("gpp", "nee", "re")
      , dep_vars_lab = c(latex2exp::TeX("$GPP$"), latex2exp::TeX("$NEE$"), latex2exp::TeX("$R_E$"))
      , st_doy = 1
      , end_doy = 366
      , yr = 2020
      , my_cov = v[i]
      , my_cov_lab = t[i]
      , sites = c("Central Plains Experimental Range", "Santa Rita Experimental Range")
    )
  )
}
```

## Scatter Plot Matrices (SPLOMS)

```{r, warning=FALSE, message=FALSE, fig.height=8, fig.width=8}
my_corr_plot_fn <- function(my_site) {
  (
    psych::pairs.panels(
        dta_1day %>% 
        dplyr::filter(neon_site_name == my_site) %>% 
        dplyr::select(
          swc, ta, ts, ppfd_in
          , vpd
          , re
          , gpp
          , nee
        )
      , pch = "."  # 21 for color by group in line above # "." for no color
      , labels = c(
          "SWC"
          , "Air Temp."
          , "Soil Temp."
          , "PPFD"
          , "VPD"
          , latex2exp::TeX("$R_E$")
          , "GPP"
          , "NEE"
      )
      , method = "pearson" # correlation method
      , hist.col = "gray70"
      , density = TRUE  # show density plots
      , ellipses = FALSE # show correlation ellipses
      , rug = FALSE
      , stars = TRUE
      , main = my_site
    )
  )  
}

# sites
sts <- unique(dta_1day$neon_site_name)
for (i in 1:length(sts)) {
  my_corr_plot_fn(my_site = sts[i])
}
```


# Draft Assignment (posted via announcement)

See "Announcements" on Canvas from Joe titled "Updated info September 13"

## Question 1 

Using the 30 minute data. Calculate GPP for your site. 

```{r}
dta_30min <- dta_30min %>% 
  dplyr::mutate(
    gpp_calc = as.numeric(re - nee)
    , gpp_is_equal = round(gpp_calc, 2) == round(gpp, 2)
  )
# quick plot of gpp equality
dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & !is.na(gpp)
  ) %>% 
ggplot(., aes(x = gpp_is_equal)) +
  geom_bar(
    aes(fill = gpp_is_equal)
    , width = 0.7
  ) +
  geom_text(
    aes(label = scales::comma(..count..))
    , stat = "count"
    , vjust = -0.5
    , color = "black"
  ) +
  scale_fill_viridis_d(alpha = 0.8) +
  scale_y_continuous(labels = scales::comma) +
  xlab("GPP Calc. = GPP? (accuracy = .01)") +
  labs(
    title = "GPP Calc. = GPP? (accuracy = .01)"
    , subtitle = paste0(site, " (30-min data)")
  ) +
  theme_bw() +
  theme(
    legend.position = "none"
  )
```

Choose a winter and summer week in your dataset. Create a plot with both your calculated GPP and the NEON GPP. 

```{r}
# select week with full data for winter and summer
sum_wk <- dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & season == "Summer"
    & year == 2020
    & is_full_week == TRUE
  ) %>% 
  dplyr::slice_sample(n = 1)
wint_wk <- dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & season == "Winter"
    & year == 2020
    & is_full_week == TRUE
  ) %>% 
  dplyr::slice_sample(n = 1)

# plot winter week
dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & year == 2020
    & week == wint_wk$week[1]
  ) %>% 
ggplot(., aes(x = time_id)) +
  geom_point(
    aes(y = gpp, color = "GPP")
    , size = 2.5
  ) +
  geom_point(
    aes(y = gpp_calc, color = "GPP Calc.")
    , size = 0.9
  ) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  xlab("Day of Year") + 
  ylab(latex2exp::TeX("GPP $(\\mu mol \\; CO_{2} \\cdot m^{-2} \\cdot$ second $^{-1})$")) +
  labs(
    title = "Winter Week Sample: Calculated GPP vs. NEON GPP"
    , subtitle = paste0(site, " (30-min data)")
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))

# plot summer week
dta_30min %>% 
  dplyr::filter(
    neon_site_name == site
    & year == 2020
    & week == sum_wk$week[1]
  ) %>% 
ggplot(., aes(x = time_id)) +
  geom_point(
    aes(y = gpp, color = "GPP")
    , size = 2.5
  ) +
  geom_point(
    aes(y = gpp_calc, color = "GPP Calc.")
    , size = 0.9
  ) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  xlab("Day of Year") + 
  ylab(latex2exp::TeX("GPP $(\\mu mol \\; CO_{2} \\cdot m^{-2} \\cdot$ second $^{-1})$")) +
  labs(
    title = "Summer Week Sample: Calculated GPP vs. NEON GPP"
    , subtitle = paste0(site, " (30-min data)")
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))
```


### a. 

How do the calculated and NEON GPP's compare for your site? Why are they the same or different? (Note: NEON uses eddy covariance to calculate these metrics - more about that [here](https://youtu.be/CR4Anc8Mkas)).

\textcolor{blue}{The majority of data points have a calculated GPP ($GPP = R_{E} - NEE$) equal to the NEON GPP. There are `r scales::comma(dta_30min %>% dplyr::filter(neon_site_name == site & gpp_is_equal == TRUE) %>% nrow())` records for which the GPP is equal (accuracy = 0.01) and `r scales::comma(dta_30min %>% dplyr::filter(neon_site_name == site & gpp_is_equal == FALSE) %>% nrow())` records for which the GPP is not equal. In cases where the GPP is not equal, the difference is negligible and likely due to rounding due to number accuracy.}


### b. 

How do your sites GPP values vary between summer and winter? What do you think is driving these differences?

```{r}
# plot points
dta_30min %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
    & season %in% c("Summer", "Winter")
  ) %>% 
ggplot(., aes(x = time_id, y = gpp, color = season)) +
  geom_point(
    # size = 0.2
    alpha = 0.4
    , shape = "."
  ) +
  geom_smooth(
    method = "lm"
    , formula=y~1
    , se = FALSE
    , lwd = 1.1
    , linetype = "solid"
    , alpha = 1
  ) +
  facet_wrap(~neon_site_name) +
  scale_color_viridis_d(option = "turbo", direction = -1) +
  xlab("Day of Year") + 
  ylab(latex2exp::TeX("GPP $(\\mu mol \\; CO_{2} \\cdot m^{-2} \\cdot$ second $^{-1})$")) +
  labs(
    title = "2020 NEON GPP (30-min data)"
    , subtitle = paste0(site)
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
    , axis.text.x = element_text(angle = 60, hjust=1)
  ) +
  guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))



```


\textcolor{blue}{Based on 2020 data, The average 30-minute GPP for the NEON site `r site` is higher in the winter months than in the summer months. This is opposite of the pattern for GPP seen at the CPER site. At the `r site`, this pattern is likely driven by low summer precipitation and high summer temperatures to the winter.}

### c. 

CPER is a semi-arid grassland. Compare GPP values between your site and CPER - why might they be different or similar?

```{r}
# plot points
dta_30min %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
    & season %in% c("Summer", "Winter")
  ) %>% 
  dplyr::group_by(neon_site_name, season, year) %>% 
  dplyr::summarise(mean_gpp = mean(gpp, na.rm = TRUE)) %>% 
ggplot(., aes(x = season, y = mean_gpp, fill = season)) +
  geom_col(
    width = 0.7
    , alpha = 0.8
  ) +
  geom_text(
    aes(label = scales::comma(mean_gpp, accuracy = 0.01))
    , vjust = -0.3
    , color = "black"
  ) +
  facet_wrap(~neon_site_name) +
  scale_fill_viridis_d(option = "turbo", direction = -1) +
  xlab("Season") + 
  ylab(latex2exp::TeX("Mean 30-min GPP $(\\mu mol \\; CO_{2} \\cdot m^{-2} \\cdot$ second $^{-1})$")) +
  labs(
    title = "2020 Seasonal NEON GPP (30-min data)"
  ) +
  theme_bw() +
  theme(
    legend.position = "none"
  )

```

\textcolor{blue}{Based on 2020 data, The average 30-minute GPP for the NEON site `r site` is higher in the winter months (`r scales::comma(mean((dta_30min %>% dplyr::filter(neon_site_name == site & year == 2020 & season %in% c("Winter")))$gpp), accuracy = 0.01)`) than in the summer months (`r scales::comma(mean((dta_30min %>% dplyr::filter(neon_site_name == site & year == 2020 & season %in% c("Summer")))$gpp), accuracy = 0.01)`). This is opposite of the pattern for GPP seen at the CPER site. At the `r site`, this pattern is likely driven by low summer precipitation and high summer temperatures to the winter.}


## Question 2 

Using the daily data. Select a single year of data for your site. Plot NEE, GPP, Re, soil temperature, air temperature, and soil water content against Day of Year.

```{r, fig.height = 4, fig.width = 7}
p_fn <- function(my_var, my_year) {
  #plot
  (
    dta_1day %>% 
    dplyr::filter(
      year == my_year
    ) %>% 
    ggplot(., aes_string(x = "date_id", y = my_var, color = "neon_site_name")) +
      geom_point() +
      xlab("Date") + 
      ylab(t[i]) +
      labs(
        title = t[i]
        , subtitle = my_year
      ) +
      scale_fill_brewer(type = "qual", palette = "Dark2", direction = -1) +
      theme_bw() +
      theme(
        legend.position = "top"
        , legend.direction = "horizontal"
        , legend.title = element_blank()
      ) +
      guides(color = guide_legend(override.aes = list(size=5, alpha = 0.9)))
  )
  
}
# vars
v <- c("nee", "gpp", "re", "ts", "ta", "swc")
t <- c(
  latex2exp::TeX("NEE $( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1})$")
   , latex2exp::TeX("GPP $( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1})$")
   , latex2exp::TeX("$R_E \\;( mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1})$")
   , "Soil Temperature (\u00B0C)"
   , "Air temperature (\u00B0C)"
    , "Soil Water Content"
)
for (i in 1:length(v)) {
  print( p_fn(my_var = v[i], my_year = 2020) )
}

```


### a. 

Describe the annual patterns in each plot and what you think drives them.

### b. 

Compare annual values of NEE at your site vs. CPER. Which exchanges more carbon? Why do you think that is?

```{r}
# plot
dta_1day %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
    & season %in% c("Summer", "Winter")
  ) %>% 
  dplyr::group_by(neon_site_name, season, year) %>% 
  dplyr::summarise(mean_nee = mean(nee, na.rm = TRUE)) %>% 
ggplot(., aes(x = season, y = mean_nee, fill = season)) +
  geom_col(
    width = 0.7
    , alpha = 0.8
  ) +
  geom_text(
    aes(label = scales::comma(mean_nee, accuracy = 0.01))
    , vjust = -0.3
    , color = "black"
  ) +
  facet_wrap(~neon_site_name) +
  scale_fill_viridis_d(option = "turbo", direction = -1) +
  xlab("Season") + 
  ylab(latex2exp::TeX("Mean daily $NEE \\; (mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1})$")) +
  labs(
    title = "2020 Seasonal NEON GPP (daily data)"
  ) +
  theme_bw() +
  theme(
    legend.position = "none"
  )

```

## Extra

Compare annual time-trend of $GPP$, $NEE$, $R_E$ for sites:

```{r}
# plot daily gpp, nee, re
dta_1day %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
  ) %>% 
ggplot(., aes(x = date_id)) +
  geom_line(
    aes(y = gpp, color = "GPP")
    , lwd = 1.2
    , alpha = 0.8
  ) +
  geom_line(
    aes(y = nee, color = "NEE")
    , lwd = 1.2
    , alpha = 0.8
  ) + 
  geom_line(
    aes(y = re, color = "RE")
    , lwd = 1.2
    , alpha = 0.8
  ) + 
  facet_wrap(~neon_site_name) +
  scale_color_brewer(type = "qual", palette = "Paired", direction = -1) +
  xlab("Date") + 
  ylab(latex2exp::TeX("$mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1}$")) +
  labs(
    title = "2020 NEON Site Carbon Fluxes (1-day data)"
  ) +
  theme_bw() +
  theme(
    legend.position = "top"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
    , axis.text.x = element_text(angle = 60, hjust=1)
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))

```

```{r}
# plot daily gpp, nee, re
dta_1day %>% 
  dplyr::filter(
    # neon_site_name == site
    year == 2020
  ) %>% 
ggplot(., aes(x = date_id)) +
  geom_smooth(
    aes(y = gpp, color = "GPP")
    , method = "loess"
    , span = 0.7
    , se = FALSE
    , lwd = 1.2
    , alpha = 0.8
  ) +
  geom_smooth(
    aes(y = nee, color = "NEE")
    , method = "loess"
    , span = 0.7
    , se = FALSE
    , lwd = 1.2
    , alpha = 0.8
  ) + 
  geom_smooth(
    aes(y = re, color = "RE")
    , method = "loess"
    , span = 0.7
    , se = FALSE
    , lwd = 1.2
    , alpha = 0.8
  ) + 
  facet_wrap(~neon_site_name) +
  scale_color_brewer(type = "qual", palette = "Paired", direction = -1) +
  xlab("Date") + 
  ylab(latex2exp::TeX("$mol \\; CO_{2} \\cdot m^{-2} \\cdot day^{-1}$")) +
  labs(
    title = "2020 NEON Site Carbon Fluxes (1-day data)"
    , subtitle = "Smoothed Data"
  ) +
  theme_bw() +
  theme(
    legend.position = "top"
    , legend.direction = "horizontal"
    , legend.title = element_blank()
    , axis.text.x = element_text(angle = 60, hjust=1)
  ) +
  guides(color = guide_legend(override.aes = list(size = 5)))

```


