---
title: "NEON_Assign1_script"
output:
  pdf_document: default
  html_document: default
date: "2022-08-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the R script for ECOL610 NEON assignment 1
This script will provide code to carry out assignment 1 for the Central Plains Experiemntal Range - you will need to modify this code to carry it out for your site.

First, load in the needed packages. Install the packages if needed.
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```


Load in the data. You will need to change the file path for your directory.

```{r}
#clear global environment
rm(list=ls())

#load in the data
CPER_30 <- read.csv("C:/Users/katie/OneDrive - Colostate/Documents/CSU/PhD/Fall 2022/Ecosystem Ecology Grad/NEON project/Final half-hourly files/Central Plains Experimental Range - 30 min.csv")
CPER_daily <- read.csv("C:/Users/katie/OneDrive - Colostate/Documents/CSU/PhD/Fall 2022/Ecosystem Ecology Grad/NEON project/Final daily files/Central Plains Experimental Range - daily.csv")
```

Question 1: Calculate GPP and compare to NEON GPP for summer and winter week
```{r}
#calculate GPP from NEE and Re using equation NEE = Re - GPP (negative values land C storage)
CPER_30$GPP_calc <- CPER_30$RE - CPER_30$NEE 

#clipping the data to a winter and summer week
 #choose your own weeks (without data gaps) for your site!
#February 16-22, 2020
CPER_winter_week <- filter(CPER_30, Year == "2020" & DOY == 47:53)
#June 21-27, 2020
CPER_summer_week <- filter(CPER_30, Year == "2020" & DOY == 173:180)

#creating long format for plotting
CWW_stacked <-gather(CPER_winter_week, "GPP_type", "GPP_value", c(9,15))
CSW_stacked <-gather(CPER_summer_week, "GPP_type", "GPP_value", c(9,15))

#winter
#plot of NEON vs calculated GPP
ggplot(CWW_stacked,aes(x=DOY.total,y=GPP_value, color = GPP_type)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(expression(paste("GPP (mol CO"[2]*" m"^{-2}, "day"^{-2}*")"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Winter GPP Comparison")

#summer
#plot of NEON vs calculated GPP
ggplot(CSW_stacked,aes(x=DOY.total,y=GPP_value, color = GPP_type)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(expression(paste("GPP (mol CO"[2]*" m"^{-2}, "day"^{-2}*")"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Summer GPP Comparison")

```
Question 2: annual patterns in C exchange and environmental properties
```{r}
#reduce data to just 2020
CPER_2020 <- filter(CPER_daily, Year == "2020")
#plot of NEE
ggplot(CPER_2020,aes(x=DOY,y=NEE)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(expression(paste("NEE (mol CO"[2]*" m"^{-2}, "day"^{-2}*")"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Annual NEE")
#plot of GPP
ggplot(CPER_2020,aes(x=DOY,y=GPP)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(expression(paste("GPP (mol CO"[2]*" m"^{-2}, "day"^{-2}*")"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Annual GPP")
#plot of Re
ggplot(CPER_2020,aes(x=DOY,y=RE)) + geom_point() + 
    xlab("Day of Year") + 
    ylab(expression(paste("Re (mol CO"[2]*" m"^{-2}, "day"^{-2}*")"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Annual Re")
#plot of soil temp
ggplot(CPER_2020,aes(x=DOY,y=TS)) + geom_point() + 
    xlab("Day of Year") + 
    ylab("Soil temperature (\u00B0C)") +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Annual Soil Temperature")
#plot of air temp
ggplot(CPER_2020,aes(x=DOY,y=TA)) + geom_point() + 
    xlab("Day of Year") + 
    ylab("Air temperature (\u00B0C)") +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Annual Air Temperature")
#plot of soil moisture
ggplot(CPER_2020,aes(x=DOY,y=SWC)) + geom_point() + 
    xlab("Day of Year") + 
    ylab("Soil Water Content (%)") +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Annual Soil Water Content")

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
